---
output:
  word_document: default
  html_document: default
---
```r
---
title: "Exploratory Data Analysis for Time-Series Anomaly Detection"
output: html_document
---

```{r setup, include=FALSE}
library(ggplot2)
library(dplyr)
library(tidyr)
library(lubridate)
library(forecast)
library(zoo)
library(plotly)
library(corrplot)
library(tseries)
library(imputeTS)
```

## Data Loading and Preprocessing

```{r load-data}
# Load the processed time-series data
data <- read.csv("data/processed/processed_time_series.csv")

# Preview the data
head(data)
summary(data)

# Convert the date column to Date format
data$date <- as.Date(data$date)

# Check for missing values
sum(is.na(data))

# Impute missing values using linear interpolation
data$value <- na.interpolation(data$value)

# Convert to time-series object
data_ts <- ts(data$value, start = c(YYYY, MM), frequency = 12)
```

## Time-Series Overview

```{r overview}
# Plot the time-series data
ggplot(data, aes(x = date, y = value)) +
  geom_line() +
  labs(title = "Time-Series Data Overview", x = "Date", y = "Value") +
  theme_minimal()

# Plot with plotly for interactivity
p <- ggplot(data, aes(x = date, y = value)) +
  geom_line(color = "blue") +
  labs(title = "Interactive Time-Series Data", x = "Date", y = "Value") +
  theme_minimal()

ggplotly(p)
```

## Handling Outliers

```{r outliers}
# Identify outliers using boxplot
boxplot(data$value, main = "Boxplot for Outlier Detection", ylab = "Value")

# Detect outliers using Z-score
data <- data %>%
  mutate(z_score = (value - mean(value)) / sd(value))

# Plot data with Z-score to identify outliers
ggplot(data, aes(x = date, y = z_score)) +
  geom_line() +
  geom_hline(yintercept = c(-3, 3), linetype = "dashed", color = "red") +
  labs(title = "Outlier Detection Using Z-Score", x = "Date", y = "Z-Score")
```

## Decomposition of Time-Series

```{r decomposition}
# Decompose the time series to see trend, seasonality, and noise
decomp <- decompose(data_ts)

# Plot the decomposition
plot(decomp)

# Seasonal and trend decomposition using LOESS
stl_decomp <- stl(data_ts, s.window = "periodic")
autoplot(stl_decomp)

# Save the decomposition plot as a file
ggsave("plots/seasonal_decomposition.png")
```

## Seasonality and Trend Analysis

```{r seasonality-trend}
# Seasonal plot to check for recurring patterns
ggseasonplot(data_ts) +
  ggtitle("Seasonal Plot of Time-Series Data")

# Subseries plot for detailed seasonal analysis
ggsubseriesplot(data_ts) + 
  ggtitle("Subseries Plot of Time-Series Data")

# Autocorrelation and partial autocorrelation plots
ggAcf(data_ts) + ggtitle("Autocorrelation Function")
ggPacf(data_ts) + ggtitle("Partial Autocorrelation Function")
```

## Rolling Statistics

```{r rolling-stats}
# Calculate rolling statistics (mean and variance)
data <- data %>%
  mutate(rolling_mean = rollmean(value, 12, fill = NA),
         rolling_sd = rollapply(value, 12, sd, fill = NA))

# Plot rolling mean and standard deviation
ggplot(data, aes(x = date)) +
  geom_line(aes(y = rolling_mean, color = "Rolling Mean")) +
  geom_line(aes(y = rolling_sd, color = "Rolling SD")) +
  labs(title = "Rolling Mean and Standard Deviation", x = "Date", y = "Value") +
  theme_minimal()
```

## Correlation Analysis

```{r correlation}
# Calculate correlation matrix if there are multiple features
correlation_matrix <- cor(data %>% select_if(is.numeric), use = "complete.obs")

# Plot the correlation matrix
corrplot(correlation_matrix, method = "color", type = "upper", tl.cex = 0.8)

# Save the correlation plot
ggsave("plots/correlation_matrix.png")
```

## Stationarity Testing

```{r stationarity-test}
# Perform Augmented Dickey-Fuller Test for stationarity
adf_test <- adf.test(data_ts, alternative = "stationary")

# Display ADF test result
adf_test

# If p-value > 0.05, difference the time series to make it stationary
data_diff <- diff(data_ts)

# Plot differenced time series
autoplot(data_diff) +
  ggtitle("Differenced Time-Series Data")
```

## Anomaly Detection Preparation

```{r anomaly-preparation}
# Plot the data with rolling mean to visualize anomalies
ggplot(data, aes(x = date, y = value)) +
  geom_line() +
  geom_line(aes(y = rolling_mean), color = "red") +
  labs(title = "Time-Series with Rolling Mean", x = "Date", y = "Value") +
  theme_minimal()

# Plot rolling standard deviation for potential anomaly identification
ggplot(data, aes(x = date, y = rolling_sd)) +
  geom_line(color = "purple") +
  labs(title = "Rolling Standard Deviation", x = "Date", y = "Rolling SD") +
  theme_minimal()
```

## Advanced Visualizations

```{r advanced-viz}
# Histogram to visualize data distribution
ggplot(data, aes(x = value)) +
  geom_histogram(binwidth = 5, fill = "blue", alpha = 0.7) +
  labs(title = "Distribution of Time-Series Values", x = "Value", y = "Frequency")

# Boxplot to visualize distribution across different time periods
ggplot(data, aes(x = factor(month(date)), y = value)) +
  geom_boxplot() +
  labs(title = "Monthly Distribution of Values", x = "Month", y = "Value") +
  theme_minimal()
```

## Correlation with External Factors

```{r external-factors}
# External data for comparison (temperature, financial index)
external_data <- read.csv("data/processed/external_factors.csv")

# Merge external factors with time-series data
merged_data <- merge(data, external_data, by = "date")

# Correlation analysis between time-series and external factors
cor_matrix <- cor(merged_data %>% select(value, external_factor1, external_factor2), use = "complete.obs")

# Plot correlation matrix
corrplot(cor_matrix, method = "ellipse", type = "upper", tl.cex = 0.8)

# Plot time-series alongside external factor
ggplot(merged_data, aes(x = date)) +
  geom_line(aes(y = value, color = "Time-Series")) +
  geom_line(aes(y = external_factor1, color = "External Factor 1")) +
  labs(title = "Time-Series vs External Factor 1", x = "Date", y = "Value") +
  theme_minimal()
```

## Anomaly Detection Using Statistical Methods

```{r anomaly-detection}
# Apply simple moving average to detect anomalies
data$moving_avg <- rollmean(data$value, 12, fill = NA)

# Identify anomalies as points where actual values deviate significantly from moving average
data <- data %>%
  mutate(anomaly = ifelse(abs(value - moving_avg) > 2 * sd(value), 1, 0))

# Plot anomalies on the time series
ggplot(data, aes(x = date, y = value)) +
  geom_line() +
  geom_point(data = subset(data, anomaly == 1), aes(x = date, y = value), color = "red", size = 2) +
  labs(title = "Anomaly Detection Using Moving Average", x = "Date", y = "Value") +
  theme_minimal()

# Save anomaly plot
ggsave("plots/anomaly_detection.png")
```